Testing Gravity using Type Ia Supernovae Discovered by LSST

### 2018-12-14 doc/article.tex
forging on in edits

Two advances in the upcoming decade will make SNe~Ia  important probes of $f\sigma_8$.
First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields $\sigma_M \gtrsim 0.12$~mag absolute magnitude
dispersion.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
Recent studies indicate that with the right data, SNe absolute
magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.
One such SN is worth $\gtrsim 25$ galaxies with 0.4~mag absolute magnitude uncertainty.
Secondly,  ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
hundreds of thousands.  This is a sample size comparable to the number of galaxies projected by WALLABY and TAIPAN.

@@

\section{Important Elements in Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
\citet{2017ApJ...847..128H} perform their analysis in Fourier space, 
where the Fisher information matrix of a random Gaussian field with mean zero and covariance $C(k)$ parameterized by $\lambda$ is
\begin{equation}
F_{ij} = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
\frac{\partial C}{\partial \lambda_j} \right],
\end{equation}
where the covariance for the velocity-velocity correlation
\begin{equation}
C = P_{vv}(k) + \frac{\sigma^2}{n}
\label{cov:eq}
\end{equation}
is dependent on the power spectrum, noise in the velocity measurement, and the density of velocity probes
\citep{2017MNRAS.464.2517H}.  
The primary dependence on the growth of structure is through
\begin{equation}
\frac{\partial C}{\partial \lambda} = \frac{\partial P_{vv}}{\partial \lambda} = \text{constant}
\end{equation}
for the parameter choice
$\lambda=(f\sigma_8)^2$.  In the sample-variance limit, $F_{\lambda \lambda} \propto V$ whereas in the shot noise limit $F_{\lambda \lambda} \propto V n^2 \sigma^{-4}$.

The survey parameters that determine parameter precision are the volume $V$, which
in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related by $\sigma_{M} = \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
Equating LSST discovery with having a meaningful peculiar velocity, $n$ is thus a non-tunable parameter.

For surveys of fixed depth, the variance of $f\sigma_8$ 
in the sample-variance limit  
is inversely proportional to the survey solid-angle $\Omega$, whereas
in the shot-noise limit the variance is inversely proportional to $\Omega n^2 \propto N^2/\Omega$, where $n$ is the number density,
and $N$ is the total number of supernovae.

\subsection{Solid Angle}

The variance in $f\sigma_8$ is proportional to solid angle $\Omega$ given a fixed redshift depth and  $\sigma^2_M/n$.  The baseline
LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
coverage of the southern equatorial pole, could double
the sky coverage and halve the variance in $f\sigma_8$.

\subsection{Intrinsic Magnitude Dispersion}
The quality of supernova follow-up is an important factor in how well a peculiar-velocity survey can measure $f\sigma_8$.
Supplemental non-LSST data can access SN models that yield lower intrinsic-magnitude dispersion and systematic errors.
LSST SN~Ia light curves from the WFD survey, with their relatively poor sampling, plus a spectroscopic redshift
 can give $\sigma_M \sim 0.15$~mag.
 Non-LSST follow-up including infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
 \citep{2015ApJ...815...58F} are projected to give
$\sigma_M \lesssim 0.08$~mag.
Uncertainties in $f\sigma_8$ depend on $\sigma^2_M/n$ so the improvement between 0.08 and 0.15~mag dispersions is equivalent to a factor of 3.52
 in number density, or in survey duration. 
While LSST data can be used to obtain photometric redshifts from host galaxies and supernova light curves, the spectroscopic
redshifts needed for precise host-galaxy peculiar velocities require other resources.

LSST can be taken to be a SN~Ia discovery plus light-curve builder, or a discovery machine only.  The former comes with a simplified follow-up program of typing
and/or redshift collection but with higher intrinsic magnitude dispersion.  The latter requires additional follow-up resources to gather more information per supernova to
lower the intrinsic magnitude dispersion.

The nominal observing strategy will obtain photometry every couple of days rotating through the 6 $grizy$ bands.  Although SNe have good
overall sampling, the per-band light curves are poorly sampled.

\appendix
\section{Reanalysis in Configuration Space}

This Appendix describes a new analysis used to project precisions on $f\sigma_8$ using peculiar velocities derived from SNe~Ia.
This analysis may or may not be useful for the Decadal Survey White Paper, but will be helpful for LSST planning.
While there have been a number of articles on the subject,
our analysis brings a higher level of fidelity than sought previously.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
catalog. The numbers of SNe are sufficiently small to allow fast evaluations of the likelihood, which enable the determination of parameter
posteriors using MCMC on reasonable computing timescales.   We can use our machinery to 
compare different survey parameters, such as redshift depth, total numbers of supernovae, non-uniform spatial density,
solid angle/survey geometry, and SN~Ia intrinsic magnitude dispersion.
Adding the cross-correlation between galaxy-count and peculiar-velocity surveys is active work, so we can
quantify the suppression of sample variance achieved when considering matter-densities and velocities 
within the same volume \citep{2007PhRvL..99h1301G}.

\section{Simulated Data}
The cosmoDC2 (v1.0) covering 706 sq.~deg is used\footnote{We previously used
Buzzard (v1.6) galaxy catalog is used, but it was found to give peculiar velocities inconsistent with general relativity.  As we will see,
cosmoDC2 (v1.0) is consistent with GR. }. 
The catalog is based on a Flat $\Lambda$CDM model with $H_0=71$~km~s$^{-1}$,  $\Omega_M=0.265$, $\Omega_B=0.0448$, and
$\Omega_\nu=0$.
Each galaxy has its cosmological redshift, the $x$-, $y$-, $z$-components of its peculiar velocity from which
the radial peculiar velocity is determined, and its
star formation rate and stellar mass from which the
supernova rate of each host is determined using 
\citet{2012ApJ...755...61S}.   This host-based rate
underestimates the expected discovery based on the volumetric
rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so  we boost the SN production
by a factor 2.4 to get the 10-year volumetric rate expectation. The total number of simulated SNe 
is greater than the total number of possibly useful supernova discoveries, since most fields cannot be continuously monitored throughout the year
from most ground-based observatories.

Each supernova is assigned a magnitude based on the distance modulus of its cosmological
redshift plus a random term drawn from a Normal distribution $\sigma_M$, which for simplicity is the same for all supernovae and captures both
intrinsic magnitude dispersion and measurement uncertainty.  No other corrections are applied to the observed magnitude.  Dipole effects of the heliocentric
motion with respect to the CMB and of the galaxies with respect to the CMB  are ignored: referring to
\citet{2011ApJ...741...67D}, the effects in Eq.~18 are not included.  Including these effects are straightforward but is computationally expensive
and does not affect our conclusions.

We consider a maximum redshift of $z=0.2$ because over ten years the noise is not sample-variance dominated, meaning that the differential improvement
in the precision of
$f\sigma_8$ due to a lower-redshift supernova is greater than that due to a high-redshift supernova.  Following all LSST SNe~Ia below this redshift 
would likely saturate available follow-up resources.

\section{Analysis}
The analysis is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}, except as noted later.  
The peculiar magnitude is given by 
\begin{equation}
\delta_m=(m - M) - \mu(z;H_0, \Omega_M, \Omega_\Lambda).
\label{data:eqn}
\end{equation}  We introduce one free parameter 
$\mathcal{M} = M - 5\log{H_0} + \text{const}$ and do not consider the other cosmological parameters, whose effects are relatively weak at low-redshift.

The peculiar magnitude correlation
function $\xi_{\delta m \delta m}$ \citep{2011ApJ...741...67D,2015JCAP...12..033H} expected from General Relativity acting on the CMB
matter density power spectrum is calculated using CAMB \citep{Lewis:2002ah}  assuming the same
cosmological parameters used for the cosmoDC2 catalog.  Our model for the data covariance
is
\begin{equation}
C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij} + \sigma^2_{NL}(z_i;\sigma_{v})\delta_{ij},
\label{cov:eqn}
\end{equation}
where $N_i$ is the number of SNe~Ia in galaxy $i$ and $\sigma_M$ is the intrinsic SN magnitude dispersion.
The model is used to test General Relativity through deviations from $A=1$ but is not sensitive to deviations
from the linearized GR expectation of the shape of the power spectrum 
$\xi$.
Nevertheless, for
convenience we say that our model gives $f\sigma_8 = A (f\sigma_8)^{GR}$, where $(f\sigma_8)^{GR}$ is the expectation
from General Relativity. 
The final term includes extra magnitude dispersion produced by non-linear effects on velocity, parameterized with $\sigma_{v}$, such that
\begin{equation}
\sigma_{NL} = \frac{5}{\ln{10}} \frac{1+\bar{z}}{\bar{z}} \sigma_v,
\end{equation}
where $\bar{z}$ is the cosmological redshift of the host galaxy.
We note that the CAMB calculation of $\xi$ includes non-linear corrections, which actually may be a bad thing for peculiar velocities \citep{2015MNRAS.454.3920H}. 
Such subtleties are unimportant as \citet{2015JCAP...12..033H} find that their results are insensitive to whether non-linear corrections are or are not included.

For all analyses we remove supernovae below observed $z_{min}=0.01$ in order to
reduce errors made in the first-order transformation between velocity and
magnitude.  This cut removes a  small volume relative to the full survey.

The posterior of the parameters is sampled using emcee.  The likelihood is the Normal distribution with the data in Eq.~\ref{data:eqn} and
covariance matrix in Eq.~\ref{cov:eqn}.
The priors are positive and flat for $A$, flat for $M$, positive Cauchy(0.08, 0.5) for $\sigma_M$, and  positive Cauchy(0, 600 km~s$^{-1}$) for $\sigma_{v}$.

Direct tests of non-GR models for which Poisson's equation or the linear approximation
does not apply requires a model-dependent $\xi$.  We defer the analysis of such higher-fidelity models to later work.

\section{Results of Subsets}

The  ston in $f\sigma_8$ is calculated for different subsamples of the simulated data.
For each subsample  the mean  and standard deviation, $(\bar{A}, \sigma_A$) of the amplitude parameter is
determined.  The ston in $f\sigma_8$ for the subsample is thus  $\bar{A} \sigma^{-1}_A$, and that for a 18,000 sq.~deg. survey  is approximated by
$\sqrt{18000/760}\bar{A}\sigma^{-1}_A$.  The effective constraining power of a single supernova in the subset is chracterized by 
 $\bar{A} \sigma^{-1}_A N_{gal}^{-0.5}$.

Different parameters describing the input simulated data define the subsamples, 
which are used to explore the range of possible survey strategies.   The varied parameters are the redshift range, $z_{max}$, intrinsic
magnitude dispersion, $\sigma_M$, and SN number density, $n = \phi n_{10}$, where $ n_{10}$ is the  10-year total number density.  In general a given point
of sky is not monitored over the full year.  
The fit posterior for one representative subsample, with input  $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \citet{2015JCAP...12..033H, 2017JCAP...05..015H},
is found to be consistent with zero, poorly constrained, and very weakly correlated with $A$.  Table~\ref{tab:subsets} gives the results for the different subsamples.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.0.65.0.2.pkl.png}
\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $\mathcal{M}$ is the supernova
absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
to peculiar velocity.  This result is for the subset $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$.  The dotted lines represent the inputs of the supernova
part of the simulation.
\label{zmax:fig}}
\end{figure}

>>>>

Two advances in the upcoming decade will make SNe~Ia  important probes of $f\sigma_8$.
First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields $\sigma_M \gtrsim 0.12$~mag absolute magnitude
dispersion.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
Recent studies indicate that with the right data, SNe absolute
magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.
One such SN is worth $\gtrsim 25$ galaxies with 0.4~mag absolute magnitude uncertainty.
Secondly,  ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
hundreds of thousands, over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia for which good light curves can be measured. This is a sample size comparable to the number of galaxies projected by WALLABY and TAIPAN.

@@

\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
\subsection{Parameters that Determine Scientific Performance}

\citet{2017ApJ...847..128H} perform their analysis in Fourier space, 
where the Fisher information matrix of a random Gaussian field with mean zero and covariance $C(k)$ parameterized by $\lambda$ is
\begin{equation}
F_{ij} = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
\frac{\partial C}{\partial \lambda_j} \right],
\end{equation}
where the covariance for the velocity-velocity correlation
\begin{equation}
C = P_{vv}(k) + \frac{\sigma^2}{n}
\label{cov:eq}
\end{equation}
is dependent on the power spectrum, noise in the velocity measurement dominated by the intrinsic velocity dispersion $\sigma$, and the density of velocity probes
\citep{2017MNRAS.464.2517H}.  
The primary dependence on the growth of structure is through
\begin{equation}
\frac{\partial C}{\partial \lambda} = \frac{\partial P_{vv}}{\partial \lambda} = \text{constant}
\end{equation}
for the parameter choice
$\lambda=(f\sigma_8)^2$.  In the sample-variance limit, $F_{\lambda \lambda} \propto V$ whereas in the shot noise limit $F_{\lambda \lambda} \propto V n^2 \sigma^{-4}$.

The survey parameters that determine parameter precision are the volume $V$, which
in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related by $\sigma_{M} = \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.

LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
Equating LSST discovery with having a meaningful peculiar velocity, the number density is$n$ is thus a non-tunable parameter, which only increases with survey duration.
The remaining free parameters are then $\Omega$, $z_{max}$, and $\sigma_M$.

While LSST data can be used to obtain photometric redshifts from host galaxies and supernova light curves, the spectroscopic
redshifts needed for precise host-galaxy peculiar velocities require other resources.

\begin{itemize}
\item Solid Angle $\Omega$:
The variance in $f\sigma_8$ is proportional to solid angle $\Omega$.  The baseline
LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
coverage of the southern equatorial pole, could double
the sky coverage and halve the variance in $f\sigma_8$.
\item Redshift Depth $z_{max}$:
For the number densities and intrinsic magnitude dispersions relevant to LSST, the noise is not sample-variance
limited.  We have run simulations of LSST SN surveys
and calculated the signal-to-noise (STON) of $f\sigma_8$ from peculiar velocities. These simulations show, as seen in the left plot
of  Figure~\ref{notshot:fig}, that an increase in the number the number density of tracers leads to significant improvements in STON.
It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of  STON on $\sigma_M^2/n$.

\begin{figure}
\centering
\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities.  For several limiting redshift depths, the STON is plotted as a
function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
\label{notshot:fig}}
\end{figure}

The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers as shown in the
right plot of   Figure~\ref{notshot:fig}, shows that LSST is in the regime where adding one tracer at lower redshift better improves the 
STON relative to the addition of one at higher redshift.

Scientific leverage dictates that low-redshift sources are most valuable.
This aligns with observational considerations, for which closer and hence brighter objects require more modest follow-up resources.
Therefore, $z_{max}$ depends on the follow-up program as the redshift where complete follow-up saturates resources.

\item Intrinsic Magnitude Dispersion $\sigma_M$:
Just as an increase in tracer number density beyond that of a ten-year LSST survey improves the STON in the measurement
of $f\sigma_8$, as noted in the discussion on $z_{max}$, so too does a decrease in the supernova intrinsic magnitude dispersion.

The nominal LSST WFD observing plan is to have
each piece of sky observed every couple of days rotating through the 6 $grizy$ bands. The data is sufficient to have effectively
complete $z<0.3$ SN~Ia discoveries.  However,  the per-band light curves are poorly sampled, which limits the accuracy with which distances
can be derived.
LSST SN~Ia light curves from the WFD survey, with their relatively poor sampling, plus a spectroscopic redshift
 can give $\sigma_M \sim 0.15$~mag.

Supplemental non-LSST follow-up data in the form or improved temporal light-curve sampling (accurate rise and decline times), 
expanded (UV, NIR) wavelength coverage, and spectral features
can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
 \citep{2015ApJ...815...58F} are projected to give
$\sigma_M \lesssim 0.08$~mag.
Uncertainties in $f\sigma_8$ depend on $\sigma^2_M/n$ so the improvement between 0.08 and 0.15~mag dispersions is equivalent to a factor of 3.52
 in number density, or in survey duration. 
\end{itemize}

\subsection{Designing an SN~Ia Peculiar Velocity Survey}
There are different options to consider in the design of a SN~Ia peculiar velocity survey, though at this point it is difficult to assess which is optimal
or most cost effective.  While there are a variety
of recently identified indicators of SN~Ia diversity that improve upon the 2-parameter model (SALT2) currently used to determine distances, 
there is not yet an umbrella model that simultaneously captures all indicators, their correlations, and characterizes the model residuals.
Constructing such a model is a parallel
endeavor that will benefit peculiar-velocity  and other SN~Ia science.  Instead of making specific recommendations, we make several general conclusions
based upon the above findings.

\begin{itemize}
\item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.
\item Role of LSST:
LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   The nominal cadence of 
 produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone.  Leaving the
 distance determination for other telescope, LSST should go for $\Omega$.
\item Follow-up resources.  21 mag.  2--4m-class telescopes.
\end{itemize}
\section{Conclusions}
%
%
%\appendix
%\section{Reanalysis in Configuration Space}
%
%This Appendix describes a new analysis used to project precisions on $f\sigma_8$ using peculiar velocities derived from SNe~Ia.
%This analysis may or may not be useful for the Decadal Survey White Paper, but will be helpful for LSST planning.
%While there have been a number of articles on the subject,
%our analysis brings a higher level of fidelity than sought previously.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
%catalog. The numbers of SNe are sufficiently small to allow fast evaluations of the likelihood, which enable the determination of parameter
%posteriors using MCMC on reasonable computing timescales.   We can use our machinery to 
%compare different survey parameters, such as redshift depth, total numbers of supernovae, non-uniform spatial density,
%solid angle/survey geometry, and SN~Ia intrinsic magnitude dispersion.
%Adding the cross-correlation between galaxy-count and peculiar-velocity surveys is active work, so we can
%quantify the suppression of sample variance achieved when considering matter-densities and velocities 
%within the same volume \citep{2007PhRvL..99h1301G}.
%
%\section{Simulated Data}
%The cosmoDC2 (v1.0) covering 706 sq.~deg is used\footnote{We previously used
%Buzzard (v1.6) galaxy catalog is used, but it was found to give peculiar velocities inconsistent with general relativity.  As we will see,
%cosmoDC2 (v1.0) is consistent with GR. }. 
%The catalog is based on a Flat $\Lambda$CDM model with $H_0=71$~km~s$^{-1}$,  $\Omega_M=0.265$, $\Omega_B=0.0448$, and
%$\Omega_\nu=0$.
%Each galaxy has its cosmological redshift, the $x$-, $y$-, $z$-components of its peculiar velocity from which
%the radial peculiar velocity is determined, and its
%star formation rate and stellar mass from which the
%supernova rate of each host is determined using 
%\citet{2012ApJ...755...61S}.   This host-based rate
%underestimates the expected discovery based on the volumetric
%rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so  we boost the SN production
%by a factor 2.4 to get the 10-year volumetric rate expectation. The total number of simulated SNe 
%is greater than the total number of possibly useful supernova discoveries, since most fields cannot be continuously monitored throughout the year
%from most ground-based observatories.
%
%
%Each supernova is assigned a magnitude based on the distance modulus of its cosmological
%redshift plus a random term drawn from a Normal distribution $\sigma_M$, which for simplicity is the same for all supernovae and captures both
%intrinsic magnitude dispersion and measurement uncertainty.  No other corrections are applied to the observed magnitude.  Dipole effects of the heliocentric
%motion with respect to the CMB and of the galaxies with respect to the CMB  are ignored: referring to
%\citet{2011ApJ...741...67D}, the effects in Eq.~18 are not included.  Including these effects are straightforward but is computationally expensive
%and does not affect our conclusions.
%
%We consider a maximum redshift of $z=0.2$ because over ten years the noise is not sample-variance dominated, meaning that the differential improvement
%in the precision of
%$f\sigma_8$ due to a lower-redshift supernova is greater than that due to a high-redshift supernova.  Following all LSST SNe~Ia below this redshift 
%would likely saturate available follow-up resources.
%
%
%\section{Analysis}
%The analysis is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}, except as noted later.  
%The peculiar magnitude is given by 
%\begin{equation}
%\delta_m=(m - M) - \mu(z;H_0, \Omega_M, \Omega_\Lambda).
%\label{data:eqn}
%\end{equation}  We introduce one free parameter 
%$\mathcal{M} = M - 5\log{H_0} + \text{const}$ and do not consider the other cosmological parameters, whose effects are relatively weak at low-redshift.
%
%
%The peculiar magnitude correlation
%function $\xi_{\delta m \delta m}$ \citep{2011ApJ...741...67D,2015JCAP...12..033H} expected from General Relativity acting on the CMB
%matter density power spectrum is calculated using CAMB \citep{Lewis:2002ah}  assuming the same
%cosmological parameters used for the cosmoDC2 catalog.  Our model for the data covariance
%is
%\begin{equation}
%C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij} + \sigma^2_{NL}(z_i;\sigma_{v})\delta_{ij},
%\label{cov:eqn}
%\end{equation}
%where $N_i$ is the number of SNe~Ia in galaxy $i$ and $\sigma_M$ is the intrinsic SN magnitude dispersion.
%The model is used to test General Relativity through deviations from $A=1$ but is not sensitive to deviations
%from the linearized GR expectation of the shape of the power spectrum 
%$\xi$.
%Nevertheless, for
%convenience we say that our model gives $f\sigma_8 = A (f\sigma_8)^{GR}$, where $(f\sigma_8)^{GR}$ is the expectation
%from General Relativity. 
%The final term includes extra magnitude dispersion produced by non-linear effects on velocity, parameterized with $\sigma_{v}$, such that
%\begin{equation}
%\sigma_{NL} = \frac{5}{\ln{10}} \frac{1+\bar{z}}{\bar{z}} \sigma_v,
%\end{equation}
%where $\bar{z}$ is the cosmological redshift of the host galaxy.
%We note that the CAMB calculation of $\xi$ includes non-linear corrections, which actually may be a bad thing for peculiar velocities \citep{2015MNRAS.454.3920H}. 
%Such subtleties are unimportant as \citet{2015JCAP...12..033H} find that their results are insensitive to whether non-linear corrections are or are not included.
%
%For all analyses we remove supernovae below observed $z_{min}=0.01$ in order to
%reduce errors made in the first-order transformation between velocity and
%magnitude.  This cut removes a  small volume relative to the full survey.
%
%The posterior of the parameters is sampled using emcee.  The likelihood is the Normal distribution with the data in Eq.~\ref{data:eqn} and
%covariance matrix in Eq.~\ref{cov:eqn}.
%The priors are positive and flat for $A$, flat for $M$, positive Cauchy(0.08, 0.5) for $\sigma_M$, and  positive Cauchy(0, 600 km~s$^{-1}$) for $\sigma_{v}$.
%
%Direct tests of non-GR models for which Poisson's equation or the linear approximation
%does not apply requires a model-dependent $\xi$.  We defer the analysis of such higher-fidelity models to later work. 
%
%\section{Results of Subsets}
%
%The  ston in $f\sigma_8$ is calculated for different subsamples of the simulated data.
%For each subsample  the mean  and standard deviation, $(\bar{A}, \sigma_A$) of the amplitude parameter is
%determined.  The ston in $f\sigma_8$ for the subsample is thus  $\bar{A} \sigma^{-1}_A$, and that for a 18,000 sq.~deg. survey  is approximated by
%$\sqrt{18000/760}\bar{A}\sigma^{-1}_A$.  The effective constraining power of a single supernova in the subset is chracterized by 
% $\bar{A} \sigma^{-1}_A N_{gal}^{-0.5}$.
%
%Different parameters describing the input simulated data define the subsamples, 
%which are used to explore the range of possible survey strategies.   The varied parameters are the redshift range, $z_{max}$, intrinsic
%magnitude dispersion, $\sigma_M$, and SN number density, $n = \phi n_{10}$, where $ n_{10}$ is the  10-year total number density.  In general a given point
%of sky is not monitored over the full year.  
%The fit posterior for one representative subsample, with input  $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
%For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \citet{2015JCAP...12..033H, 2017JCAP...05..015H},
%is found to be consistent with zero, poorly constrained, and very weakly correlated with $A$.  Table~\ref{tab:subsets} gives the results for the different subsamples.
%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.0.65.0.2.pkl.png}
%\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $\mathcal{M}$ is the supernova
%absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
%to peculiar velocity.  This result is for the subset $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$.  The dotted lines represent the inputs of the supernova
%part of the simulation.
%\label{zmax:fig}}
%\end{figure}
%
%\begin{table}
%   \centering
%   %\topcaption{Table captions are better up top} % requires the topcapt package
%   \begin{tabular}{|ccc|cc|} % Column formatting, @{} suppresses leading/trailing space
%   \hline
%$z_{max}$ & $\phi$ & $\sigma_{M}$  &  STON $f\sigma_8$ & STON per SN \\
%\hline
%0.07 & 0.65 & 0.08  & 25.795 &   0.27 \\
%0.07 & 1.00 & 0.08  & 31.163 &   0.27 \\
%0.07 & 1.00 & 0.10  & 25.421 &   0.22 \\
%0.07 & 1.00 & 0.12  & 22.173 &   0.19 \\
%0.07 & 1.00 & 0.15  & 18.831 &   0.16 \\
%0.10 & 0.30 & 0.08  & 23.784 &   0.24 \\
%0.10 & 0.50 & 0.08  & 23.501 &   0.18 \\
%0.10 & 0.65 & 0.08  & 24.994 &   0.17 \\
%0.10 & 0.65 & 0.10  & 20.866 &   0.15 \\
%0.10 & 0.65 & 0.12  & 18.778 &   0.13 \\
%0.10 & 1.00 & 0.08  & 31.704 &   0.18 \\
%0.10 & 1.00 & 0.10  & 25.687 &   0.15 \\
%0.10 & 1.00 & 0.12  & 22.982 &   0.13 \\
%0.15 & 0.30 & 0.08  & 24.799 &   0.15 \\
%0.15 & 0.30 & 0.12  & 19.527 &   0.12 \\
%0.15 & 0.50 & 0.08  & 29.550 &   0.14 \\
%0.15 & 0.65 & 0.08  & 33.921 &   0.14 \\
%0.15 & 0.65 & 0.10  & 28.655 &   0.12 \\
%0.15 & 0.65 & 0.12  & 24.111 &   0.10 \\
%0.15 & 1.00 & 0.08  & 40.158 &   0.13 \\
%0.20 & 0.20 & 0.08  & 26.487 &   0.12 \\
%0.20 & 0.30 & 0.08  & 29.828 &   0.11 \\
%0.20 & 0.30 & 0.10  & 24.019 &   0.09 \\
%0.20 & 0.30 & 0.12  & 19.181 &   0.07 \\
%0.20 & 0.50 & 0.08  & 36.680 &   0.11 \\
%0.20 & 0.65 & 0.08  & 41.591 &   0.11 \\
%0.20 & 0.65 & 0.10  & 35.196 &   0.09 \\
%0.20 & 0.65 & 0.12  & 27.987 &   0.07 \\
%0.20 & 1.00 & 0.08  & 48.368 &   0.10 \\
%0.20 & 1.00 & 0.10  & 40.835 &   0.08 \\
%0.20 & 1.00 & 0.12  & 37.875 &   0.08 \\
%    \hline
%   \end{tabular}
%   \caption{ston in $A$ or equivalently $f\sigma_8$ for surveys parameterized by the maximum redshift $z_{max}$,
%   the fraction of total supernovae in 10 years ``fraction'', intrinsic magnitude dispersion $\sigma_{SN}$.  The number of galaxies
%   $N_{gal}$, the signal $\bar{A}$ and its uncertainty $\sigma_A$ are for the cosmoDC2 (v1.0) 760 sq.~deg. volume only.
%   The uncertainty ``per'' supernova is given by  $\sigma_A N_{gal}^{-0.5}$.  The ston, scaled to an LSST solid angle, is given as 
%    $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$.
%   \label{tab:subsets}}
%\end{table}
%
%In Fourier space, the data covariance is a combination of sample and shot noise (Eq.~\ref{cov:eq}).
%In Figure~\ref{scaling:fig} the effective LSST ston is plotted as a function of input $n \sigma^{-2}_M$ for different $z_{max}$.
%For all cases, the ston improves with larger $n \sigma^{-2}_M$ showing that the sample-variance limit has not been reached,
%though the shallowing of the slopes for shallower surveys shows that it does contribute non-negligibly.
%The fact that the ston is not constant indicates that not of the subsamples is sample-variance limited, though its non-linearity
%indicates that it is non-negligible.
%
%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\textwidth]{../outcosmo/fracsnsig2_.png}
%%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
%\caption{ Effective LSST ston as a function of input  $n \sigma^{-2}_M$.  The points are color-coded for different $z_{max}$.
%\label{scaling:fig}}
%\end{figure}
%
%A SN peculiar survey benefits from greater redshift depth $z_{max}$ due to the increase in volume and numbers of supernovae.  However,  for fixed magnitude uncertainty
%peculiar-velocity uncertainties increase with redshift and the follow-up of fainter distant SNe requires more resources. 
%Table~\ref{tab:subsets}
%shows that the effective worth of each supernova ($\sigma_A N_{gal}^{-0.5}$) diminishes with increasing survey depth.
%The table also shows that for $\sigma_M=0.08$~mag, the subsample with $z_{max}=0.1$ with 1294 supernovae yields a ston of 15.96, whereas
%a sample with $z_{max}=0.2$ with 2021 supernovae has a lower ston of 13.61.  Within the range of surveys considered, a lower-redshift supernova
%is more valuable than one at high redshift.  For both scientific and resource reasons, it is best to devote resources to the lowest possible redshifts.
%
%
%$\phi \sim 0.65$ is an effective number density of discovered supernovae with light-curve coverage observed in 10 years
%
